---
title: "Interactive Graphic Example:<br> Covid-19 Vaccinations in Suffolk County, MA"
author: "Chris Krenz"
date: "2023-03-10"
output: html_document
---

### Dataset: COVID-19 Vaccinations in the Suffolk County, Massachusetts, United States
 - Source: United States Centers for Disease Control and Prevention
 - URL: 
https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh
 - Downloaded: 3/8/2023
 - See the data dictionary in doc/DataDictionary_v36_12082022.xlsx for additional information.
  - https://github.com/chris-krenz/R-Interactive-Graphics/tree/main/doc


### R: Setup...
<details>
  <summary>Expand to see R code chunks...</summary>
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# library(shiny)
library(plotly)      # allow interactivity w/o need for shiny server hosting
library(crosstalk) 
# library(stringi)   # Can use instead of nested gsub for mult col renames
# library(lubridate) # Can use to more easily extract year from a date

```
</details>


### R: Importing...
<details>
  <summary>Expand to see R code chunks...</summary>
```{r importing}
c19vacsMA_raw <- read_csv("data/COVID-19_Vaccinations_MA_County.csv",
                    skip = 1,   # source URL
                    na = c("UNK", ""),    # only FIPS contains 'UNK'
                    col_types = cols(Date = col_date("%m/%d/%Y")))

problems(c19vacsMA_raw) # No problems (3/8/2023)

c19vacsMA_raw

```
</details>


### R: Tidying...
<details>
  <summary>Expand to see R code chunks...</summary>
```{r tidying}
# Remove CDC analyses and other fields due to insufficient or inconsistent data
# (e.g. no pop level data for bivalent, missing age groups for 2nd booster, etc)
c19vacsMA_sel <- c19vacsMA_raw %>% 
  select(!matches("State|FIPS|Second|Equity|_SVI|Completeness|Census|5to17|50Plus|Administered_Dose1_Recip$|Administered_Dose1_Pop_Pct$|Series_Complete_Yes$|Series_Complete_Pop_Pct$|Booster_Doses$|Booster_Doses_Vax_Pct$"))  

# Rename cols to more consistent scheme ( e.g. [Dose]_[Age]_[Cnt/Pct] )
col_renames <- read_csv("data/renames.csv")
names(c19vacsMA_sel) <- 
  col_renames$New_Name[match(names(c19vacsMA_sel), col_renames$Original_Name)]

# More transparent/less redundant value names
c19vacsMA_sel$SVI[c19vacsMA_sel$SVI == "A"] <- "0.00-0.25"
c19vacsMA_sel$SVI[c19vacsMA_sel$SVI == "B"] <- "0.25-0.50"
c19vacsMA_sel$SVI[c19vacsMA_sel$SVI == "C"] <- "0.50-0.75"
c19vacsMA_sel$SVI[c19vacsMA_sel$SVI == "D"] <- "0.75-1.00"
c19vacsMA_sel$County <- str_replace(c19vacsMA_sel$County, " County", "")

c19vacsMA_sel


# Pivoting (longer) to consolidate various dosages and age groups to 2 cols
c19vacsMA_piv <- c19vacsMA_sel %>% 
  pivot_longer(cols = matches("Cnt$|Pct$"), 
               # names_pattern = paste0(
               #   ".*(First|Second|Booster|Bivalent)",
               #   ".*((?<!6)5+|12+|18+|65+)$"), # (?<!) look-behind
               # Using pattern: 1st term of names_to matches 1st regex group...
               names_sep = '_',   # Instead use sep, relying on simplified names
               names_to = c("Dose", "Age", "Type"),
               values_to = "Amount") %>%
    # Easier/cleaner to first pivot long, then wide, rather than a single pivot
    pivot_wider(names_from = "Type", values_from = "Amount")

c19vacsMA_piv

```
</details> 


### R: Plotting...
<details>
  <summary>Expand to see R code chunks...</summary>
```{r plotting}
plot_spacing <- theme(plot.margin = unit(c(2, 1, 1, 1), "cm"), 
                      plot.title = element_text(hjust = 0.5, vjust = 10), 
                      legend.title.align = 0.5, 
                      axis.title.x = element_text(vjust = -5), 
                      axis.title.y = element_text(vjust = 5))

# Plot 1
fig <- c19vacsMA_piv %>%
  filter(Dose == "First", 
         County == "Suffolk") %>% 
         # format(Date, "%Y") == "2021") %>%
  group_by(Week) %>% 
  
  ggplot(aes(Date, Pct, color = Age)) +
  geom_point(alpha = 0.5) +
  labs(title = "Percentage of age groups with at least 
one vaccine dose (Suffolk County, MA)
(hover over points for details, drag to zoom)",
         # ggplotly cannot accept Expressions in plot titles...
         # expression("Percentage of age groups with" ~bolditalic
         #                  ("at least")* " one vaccine dose (Suffolk County)"),
       x = "Date", y = "Age Group Percentage", color = "Age Group
(click to toggle)") +
  scale_color_manual(breaks = c('65+', '18+',    '12+',   '5+'), 
                     values = c("red", "orange", "green", "blue")) +
  plot_spacing

plotlyfig <- ggplotly(fig)

# Fix legend order from transfer to plotly
twelve    <- plotlyfig[["x"]][["data"]][[1]]
eighteen  <- plotlyfig[["x"]][["data"]][[2]]
five      <- plotlyfig[["x"]][["data"]][[3]]
sixtyfive <- plotlyfig[["x"]][["data"]][[4]]
plotlyfig[["x"]][["data"]][[1]] <- sixtyfive
plotlyfig[["x"]][["data"]][[2]] <- eighteen
plotlyfig[["x"]][["data"]][[3]] <- twelve
plotlyfig[["x"]][["data"]][[4]] <- five

```
</details>


```{r figures, echo = FALSE}
plotlyfig

```
