---
title: "Interactive Graphic Example - Covid-19 Vaccinations in Massachusetts"
author: "Chris Krenz"
date: "2023-03-08"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Introduction
### Background

### Purpose

### Dataset: COVID-19 Vaccinations in the Massachusetts
 - Source: United States Centers for Disease Control and Prevention
 - URL: 
https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh
 - Downloaded: 3/8/2023
 
### Methods
 - Dependent Vars:
   - Vaccination counts overall and for each age group:
     - 5+
     - 12+
     - 18+
     - 65+
   - Note: Only age groups with available data for all doses are included.
   - Note: 5+ counts are used for total counts.
 - Independent Vars:
   - Date (Range: 12/13/2020 - 2/15/2023)
   - MMWR Week (Morbidity and Mortality Weekly Report) (52-53 per year)
   - County (includes all 14 MA counties and 'Unknown County')
     - Note: Counties are also identified by the FIPS code (Federal Information Processing Standard).
   - Metro Status (Metro or Non-Metro)
   - SVI_CTGY: CDC Social Vulnerability Index (SVI) Rank
     - A: 0–0.25 SVI rank
     - B: 0.2501–0.50 SVI rank
     - C: 0.5001–0.75 SVI rank
     - D: 0.7501–1.0 SVI rank

See the data dictionary in doc/DataDictionary_v36_12082022.xlsx for additional information.


### R: Importing...
```{r importing}
c19vacsMA_raw <- read_csv("data/COVID-19_Vaccinations_MA_County.csv",
                    skip = 1,   # source URL
                    na = c("UNK", ""),    # only FIPS contains 'UNK'
                    col_types = cols(Date = col_date("%m/%d/%Y")))

problems(c19vacsMA_raw) # No problems (3/8/2023)

c19vacsMA_raw
```


### R: Tidying...
```{r tidying}
c19vacsMA_sel <- c19vacsMA_raw %>% 
  select(Date, MMWR_week, Recip_County, Metro_status, SVI_CTGY,
         matches("^(Census2019_).*(5Plus|12Plus|18Plus|65Plus)"),
         matches(paste0("^(Administered_Dose1_|", # all doses except second booster...
                          "Series_Complete_|",
                          "Booster_Doses_|",
                          #"Second_Booster_|",
                          "Bivalent_Booster_|)",
                       ".*(5Plus|12Plus|18Plus|65Plus)$"))) # ... and select ages

c19vacsMA_sel

# c19vacsMA_piv <- c19vacsMA_sel %>% 
#     pivot_longer(matches("Dose1"),
#     names_to = "Dose1_age_group",
#     values_to = "Dose1_pop")

c19vacsMA_piv <- c19vacsMA_sel %>% 
    pivot_longer(cols = Census2019_5PlusPop:Bivalent_Booster_65Plus, 
                 names_pattern = ".*(Census|Dose1|Series|Booster|Bivalent).*(5Plus|12Plus|18Plus|65Plus)",
                 names_to = c("Dose", "Age"),
                 values_to = "Count")

c19vacsMA_piv$Dose[c19vacsMA_piv$Dose == "Census"] <- "Total_Pop"
c19vacsMA_piv$Dose[c19vacsMA_piv$Dose == "Dose1"]  <- "First"
c19vacsMA_piv$Dose[c19vacsMA_piv$Dose == "Series"] <- "Second"
c19vacsMA_piv$Age <- str_replace(c19vacsMA_piv$Age, "Plus", "+")
rename(c19vacsMA_piv, 
       "Week"   = "MMWR_week", 
       "County" = "Recip_County", 
       "Metro"  = "Metro_status", 
       "SVI"    = "SVI_CTGY")

#TODO: Why exactly no overall data for bivalent? Assumed to be 5yo+?
#TODO: Why exactly is there so little data on Second_Booster?
```


### R: Analyzing...
```{r analyzing}

```


### R: Plotting...
```{r plotting}
plot_spacing <- theme(plot.margin = unit(c(1, 1, 1, 1), "cm"), 
                      plot.title = element_text(hjust = 0.5, vjust = 5), 
                      legend.title.align = 0.5, 
                      axis.title.x = element_text(vjust = -5), 
                      axis.title.y = element_text(vjust = 5))

```


# Conclusion
### Discussion

### References

### Notes

### Credits

